---
title: "Yongqiang Zhou"
---

## Introduction
My name is _Yongqiang Zhou_. I’m currently pursuing my Master’s in **Quantitative Methods and Modeling**, with a strong foundation in data analytics and computer science. I specialize in building predictive models and developing data-driven solutions that don’t just explain trends but optimize outcomes.\

## Linkedin
<https://www.linkedin.com/in/yongqiang-zhou-460018335/>\


```{r}
#| include: false
1+1
```
--------------
Last Updated: `r format(Sys.time(), "%A %m %d, %Y at %H:%M%p")`

```{r}
#| echo: false
#| message: false
#| warning: false

if(!require("leaflet")){
    options(repos=c(CRAN="https://cloud.r-project.org"))
    install.packages("leaflet")
    stopifnot(require("leaflet"))
}

baruch_longitude <- -73.98333
baruch_latitude  <- +40.75028

leaflet() |>
  addTiles() |>
  setView(baruch_longitude, baruch_latitude, zoom=17) |>
  addPopups(baruch_longitude, baruch_latitude, 
            "I am a Master's student at <b>Baruch College</b>!")
```